---
title: "sa2_rmarkdown_example"
author: "Daniel Fuller"
date: "12/02/2020"
output:
      html_document:
        keep_md: true
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pastecs)
library(knitr)
```

## Reading in the data

Need to remove NA too

```{r}
data <- read_csv("exploring data.csv")

data <- data %>%
          na_if(777) %>%
          na_if(888) %>%
          na_if(999)
```

## Variables

There are 2 main variables of interest

1. `depressionT1` a continuous measure of depression. 
  - Depression is the dependent variable of this study which was measured using the 20-item Center for Epidemiologic Studies Depression Scale (CES-D; Radloff, 1977). The CES-D is a common measure of depressive symptoms. Respondents are requested to rate the extent to which they experienced 20 symptoms over the past week (eg, “I felt sad”). Items are scored on a 4-point scale, ranging from 0 (rarely or none of the time, less than a day) to 3 (most or all the time, 5–7 days). Scores are computed as the sum of the responses to the items and total scores can range from 0 to 60, with higher scores indicating greater depression symptoms (Time 1: cesd101-cesd120). Scores of 16 or above indicate probable cases of clinical depression.

2. `income1` an ordinal measure of income catories. 
  - Income is the independent variable in the study. Income was measured from <$5,000 to >$40,000 in $5,000 increments. There were 9 income categories. 

### Checking variables just too make sure I have what I think I do. 

```{r}
typeof(data$depressionT1)
typeof(data$income1)
#data$income1 <- as.integer(data$income1)
#typeof(data$income1)
```

`income1` is type double (a number) not an ordinal variable. We need to recode this to ordinal. 

```{r}
data$income <- as.ordered(data$income1)
typeof(data$income)
```

## Sample Description
```{r}
table(data$cgsex)
sum(table(data$cgsex))

summary(data$educ)

summary(data$income1)

table(data$mstat1)

table(data$dement1)
```

At Time 1, 77.1% of the caregivers were female. Mean caregiver age was 62.8 (SD = 14.23), and the sample was predominately Caucasian (79.0%). The caregivers in this sample were highly educated, with 60.3% attaining at least some university or college education. Median household income was $25,000 to 30,000 per year (SD = $15,000). Almost half of caregivers were spouses (45.8%) and more than one third (38.7%) were adult children of care recipients. Care recipients had a wide variety of disabling conditions with 41% diagnosed with Alzheimer’s disease or other forms of dementia.

Figures 1 and 2 are histograms for the independent and dependent variables. The histogram for activity restriction in sports and recreation appears to be somewhat normally distributed, but has a positive skew due to the mode of this data being 1 (on a 5-point scale 0-4). The histogram for depression illustrates that the data is very positively skewed and there appears to be several outliers who have very severe depression.

## Descriptive statistics for `depressionT1`

Using functin `stat.desc` from `pastecs` package combined with kable from the `knitr` to make a nice table

```{r}
kable(stat.desc(data$depressionT1))
```

## Descriptive statistics for `income1`

Income1 is ordinal so doing descriptives like mean is not meaningful. Instead I use frequency tables. For full marks you should recode this to it's meaningful when displayed. 

```{r}
count <- table(data$income)  # Table of counts
kable(count, format = "markdown") #Display
```

## Histogram and boxplot for `depressionT1`

```{r}
histo_dep <- ggplot(data, aes(depressionT1)) + 
                geom_histogram() + 
                ggtitle("Figure 1. Histogram of Depression")
plot(histo_dep)

box_dep <- ggplot(data, aes("depressionT1", depressionT1)) + 
                geom_boxplot() + 
                ggtitle("Figure 2. Boxplot of Depression")
plot(box_dep)
```

## Kolmogorov-Smirnov test

K-S test function includes 2 inputs. First, your variable, second the distribution you want to compare your distribution against. Here we put `rnorm` because we can to compare our distribution against a normal distribution.

```{r}
ks.test(data$depressionT1, rnorm)
```

## Boxplot for `income1`

```{r}
bar_income <- ggplot(data, aes(income1)) + 
                geom_bar() + 
                ggtitle("Figure 3. Barplot of Income Categories")
plot(bar_income)
```

## Normality Tests

Tests for normality were performed. Depression was extremely positive skewed (Zskewness = 8.40) and the distribution was leptkurtic (Zkurtosis = 3.56). Depression at time 1 was not  (D = 3.77, p < .001). Box plots indicated six outliers in depression; these were caregivers who had high depression scores (score of 38 and over). On average caregivers were slightly restricted in their sports and recreation activates due to


## Scatter plot between `depressionT1` and `income1`

```{r}
scatter_inc_dep <- ggplot(data, aes(income1, depressionT1)) + 
                geom_point() + 
                ggtitle("Figure 4. Scatterplot of Income Categories\nand Depression")
plot(scatter_inc_dep)
```

## Correlation 

Choosing Spearman correlation but we need to use `income1` the integer variable to get the correlation value. 

```{r}
cor.test(data$depressionT1, data$income1, method = "spearman")
```

### Correlations

Figure 4 illustrates a scatter plot of income and depression. This graph illustrates that there does not generally appear to be a linear relationship between income and depression. However, it does appear that in general, individuals with greater income did experience less depressive symptoms. 

A bivariate correlation was calculated between income and depression. Spearman’s rho correlation coefficient was chosen as the variables are not normally distributed and included an ordinal variable (income) (Field, 2012). These variables were negative correlated (rho = .105, p = .001) indicating that there was a weak relationship between income and levels of depression.

