---
title: "sa6_rmarkdown_example"
author: "Daniel Fuller"
date: "06/03/2020"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Repeated Measures ANOVA

### Checking the data

```{r}
library(tidyverse)
library(car)
library(kableExtra)
library(rstatix)

iron_data <- read_csv("https://raw.githubusercontent.com/walkabilly/HKR_6000/master/assignments/assignment6/RM_ANOVA.csv")

str(iron_data$id)
iron_data$participant_id <- as.factor(iron_data$id)
str(iron_data$id)

str(iron_data$group)
str(iron_data$dose)
str(iron_data$iron_level)
```

### Changing the level order 

```{r}
iron_data$group <- fct_relevel(iron_data$group, c("no vitamin C", "vitamin C"))
iron_data$dose <- fct_relevel(iron_data$dose, c("verylow", "low", "medium", "high"))
```

### Descriptive statistics

```{r}
summary(iron_data$iron_level)

desc_group <- iron_data %>%
                group_by(group, dose) %>%
                  summarise(
                    mean_group = mean(iron_level, na.rm = TRUE),
                    sd_group = sd(iron_level, na.rm = TRUE),
                    median_group = median(iron_level, na.rm = TRUE),
                    variance_group = var(iron_level, na.rm = TRUE),
                  )
kable(desc_group, format = "markdown")
```

### Histogram
```{r}
iron_histogram <- ggplot(iron_data, aes(iron_level)) + 
                  geom_histogram()
plot(iron_histogram)
```

### Histogram by group

```{r}
iron_dose_histogram <- ggplot(iron_data, aes(iron_level)) + 
                  geom_histogram() + 
                  facet_wrap(~ dose)
plot(iron_dose_histogram)
```

### Histogram by group

```{r}
iron_group_histogram <- ggplot(iron_data, aes(iron_level)) + 
                  geom_histogram() + 
                  facet_wrap(~ group)
plot(iron_group_histogram)
```

### Scatter plot iron by group by dose

```{r}
group_iron <- ggplot(iron_data, aes(x = dose, y = iron_level, fill = group)) + 
                  geom_boxplot() 
plot(group_iron)
```

### Adjusted scatter plot iron by group by dose

```{r}
iron_data <- iron_data %>%
                  group_by(id) %>%
                    mutate(
                      id_mean = mean(iron_level, na.rm = TRUE)
                    )

iron_data$grand_mean <- mean(iron_data$iron_level)

iron_data$adj <- iron_data$grand_mean - iron_data$id_mean

iron_data$adj_iron_data <- iron_data$adj + iron_data$iron_level
```

```{r}
group_iron_adj <- ggplot(iron_data, aes(x = dose, y = adj_iron_data, fill = group)) + 
                  geom_boxplot() 
plot(group_iron_adj)
```

### Repeated Measures ANOVA analysis - Planned Contrasts

Actually, we only have two groups so we don't need fancy contrasts here. 

```{r}
### We will adjust the error term in the ANOVA but basically it's the same calls as Assignment 5. 

res.aov <- anova_test(data = iron_data, dv = iron_level, wid = participant_id, between = group, within = dose)
get_anova_table(res.aov)


### Run the anova
options(contrasts = c("contr.sum","contr.poly"))
anova_1 <- aov(iron_level ~ group + Error(id/group), data = iron_data)
summary(anova_1) 


results <- get_contrasts(anova_1)
print(results$contrasts)
```

### ANOVA analysis - Polynomial contrasts

```{r}
### Us the function get to polynomial contrats
contrasts(stress_data$group) <- contr.poly(3)
anova_2 <- aov(stress ~ group, data = stress_data)

summary.aov(anova_2, split=list(group=list("linear"=1, "quadratic" = 2))) 
```

### Tukey posthoc test

```{r}
tukey_hsd_test<- PostHocTest(anova_1, method = "hsd")
tukey_hsd_test

bonf_test <- PostHocTest(anova_1, method = "bonferroni")
bonf_test
```

### Games Howell

```{r}
pairwise_comp <- pairwise_comparisons(data = stress_data, x = group, y = stress, type = "parametric", var.equal = FALSE )
pairwise_comp
```


